---
title: "afcestDataCheck: Comprehensive Reference Manual"
subtitle: "High-Frequency Data Quality Checks for Survey Research -- Version 0.3.0"
author: "Cedric Bationo, African Center for Studies and Training (AfCEST)"
date: "`r Sys.Date()`"
output:
  rmarkdown::pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
    fig_caption: true
    latex_engine: xelatex
    keep_tex: false
fontsize: 11pt
geometry: margin=1in
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{xcolor}
  - \usepackage{fancyhdr}
  - \usepackage{enumitem}
  - \usepackage{float}
  - \definecolor{darkblue}{RGB}{0,51,102}
  - \definecolor{accentgreen}{RGB}{39,119,42}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \fancyhead[L]{\small\textit{afcestDataCheck}}
  - \fancyhead[R]{\small\textit{Reference Manual}}
  - \fancyfoot[C]{\thepage}
  - \renewcommand{\headrulewidth}{0.4pt}
vignette: >
  %\VignetteIndexEntry{afcestDataCheck: Comprehensive Reference Manual}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  eval      = FALSE,
  echo      = TRUE,
  collapse  = TRUE,
  comment   = "#>",
  fig.align = "center"
)
```

\vspace{1cm}

\begin{abstract}
\noindent
\textbf{afcestDataCheck} is an R package providing 65+ automated
high-frequency data quality check (HFC) functions for survey research
in international development. The package covers 13 categories of
checks---from duplicate identification and GPS validation to
fabrication detection and back-check reconciliation---and supports
five major data collection platforms (KoBoToolbox, SurveyCTO,
ODK Central, Survey Solutions, and CSPro). All checks return a
standardized \texttt{check\_result} object, enabling seamless
orchestration via YAML-driven configuration files and one-command
export to formatted Excel workbooks. This vignette serves as the
authoritative reference manual for the package.
\end{abstract}

\newpage

# Introduction

## Motivation

Field surveys are the backbone of evidence-based policy in
international development. Yet data quality problems---duplicate
records, fabricated responses, GPS anomalies, enumerator
misconduct---frequently compromise survey integrity. Traditional
quality assurance relies on ad-hoc Stata do-files or manual spot
checks, neither of which scales to the volume and velocity of modern
mobile data collection.

**afcestDataCheck** addresses this gap by providing a
comprehensive, modular R toolkit for automated high-frequency checks
(HFCs). The package is designed with three guiding principles:

\begin{enumerate}[leftmargin=2cm]
  \item \textbf{Completeness.} 65+ check functions organized into
        13 thematic categories cover the full HFC lifecycle, from
        sample tracking through back-check reconciliation.
  \item \textbf{Standardization.} Every check function returns a
        uniform \texttt{check\_result} list, making it trivial to
        aggregate, filter, and export results programmatically.
  \item \textbf{Configurability.} A YAML-driven configuration system
        lets survey managers define project-specific parameters once
        and run the entire suite with a single call to
        \texttt{run\_all\_checks()}.
\end{enumerate}

## Scope

The package targets development-sector surveys including, but not
limited to, agricultural production surveys, household expenditure
surveys, health facility assessments (SDI), and food security
monitoring (CFSVA, EFSA). It integrates natively with KoBoToolbox,
SurveyCTO, ODK Central, Survey Solutions, and CSPro through a
platform-normalization layer.

## How to Read This Manual

Sections 2--4 provide a quick-start guide. Sections 5.1--5.13
constitute the comprehensive function reference, organized by
category. Sections 6--10 describe the configuration system,
orchestrator, platform normalization, and export facilities.
Section 12 contains the complete function index as a reference
table.

\newpage

# Installation

## From GitHub (recommended)

```{r install-github}
# install.packages("pak")
pak::pak("afcest/afcestDataCheck")
```

## From source

```{r install-source}
# Clone the repository, then:
devtools::install("path/to/survey-data-quality")
```

## Optional dependencies

Several check functions use optional packages. Install them as
needed:
```{r install-optional}
# Spatial checks (GPS boundary, clustering)
install.packages("sf")

# Excel export with conditional formatting
install.packages("openxlsx2")

# Platform API access
install.packages(c("robotoolbox", "rsurveycto", "ruODK"))
```

\newpage

# Quick Start

This section demonstrates the core workflow: load data, run checks,
inspect results, and export.

## Step 1: Load the bundled dataset

```{r quickstart-data}
library(afcestDataCheck)

# Load the bundled dataset (1,206 rows from real World Bank iehfc survey data)
data(survey_sample)

# Preview
head(survey_sample[, c("hh_id", "enum_id", "submission_date", "duration",
                        "income", "gps_lat", "gps_lon")])
```

## Step 2: Run individual checks

```{r quickstart-individual}
# Check for duplicate IDs
res_dup <- check_duplicate_ids(survey_sample, id_col = "hh_id")
print(res_dup)
#> -- check_result: duplicate_hh_id ----
#>   Severity : error
#>   Flagged  : 6 observation(s)
#>   IDs      : 5045, 5045, 5067, 5067, 5098, 5098

# Check survey duration
res_dur <- check_survey_duration(
  survey_sample, id_col = "hh_id", duration_col = "duration",
  min_duration = 15, max_duration = 120
)
print(res_dur)

# Check GPS null island
res_gps <- check_gps_null_island(
  survey_sample, id_col = "hh_id", lat_col = "gps_lat", lon_col = "gps_lon"
)
print(res_gps)
```

## Step 3: Run all checks via configuration

```{r quickstart-config}
# Load the bundled example configuration (matches survey_sample column names)
config <- read_config(
  system.file("example_config.yml", package = "afcestDataCheck")
)
report <- run_all_checks(survey_sample, config = config, verbose = TRUE)
print(report)
summary(report)
```

## Step 4: Export results

```{r quickstart-export}
# Export to a formatted Excel workbook
export_to_excel(report, data = survey_sample, id_col = "hh_id",
                path = "hfc_report.xlsx",
                enum_col = "enum_id", date_col = "submission_date")

# Or export to CSV
export_to_csv(report, output_dir = "hfc_output/",
              data = survey_sample, id_col = "hh_id",
              enum_col = "enum_id", date_col = "submission_date")
```

\newpage

# The `check_result` Object

Before diving into the function reference, it is essential to
understand the data structure that every check function returns.

## Structure

Every check function returns a named list of class
\texttt{check\_result} with the following fields:

\begin{table}[H]
\centering
\caption{Fields of a \texttt{check\_result} object}
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Description} \\
\midrule
\texttt{check\_name}     & \texttt{character(1)} & Machine-readable check identifier \\
\texttt{check\_category} & \texttt{character(1)} & Category (e.g., ``identification'', ``enumerator'') \\
\texttt{n\_flagged}      & \texttt{integer(1)}   & Number of observations flagged \\
\texttt{n\_total}        & \texttt{integer(1)}   & Total number of records examined \\
\texttt{flagged\_ids}    & \texttt{character()}  & Vector of IDs that were flagged \\
\texttt{flag\_reason}    & \texttt{character()}  & Human-readable reason per flag \\
\texttt{severity}        & \texttt{character(1)} & One of \texttt{"error"}, \texttt{"warning"}, \texttt{"info"} \\
\texttt{summary\_stat}   & \texttt{list()}       & Check-specific summary statistics \\
\texttt{timestamp}       & \texttt{POSIXct}      & Timestamp when the check was run \\
\bottomrule
\end{tabular}
\end{table}

## Constructor and helpers

```{r check-result-helpers}
# Create a check_result manually (used internally)
res <- new_check_result(
  check_name     = "my_custom_check",
  check_category = "logic",
  n_flagged      = 3L,
  n_total        = 200L,
  flagged_ids    = c("HH-001", "HH-007", "HH-042"),
  flag_reason    = rep("Custom logic violation", 3),
  severity       = "warning",
  summary_stat   = list(total_obs = 200, pct_flagged = 1.5)
)

# Validate
is_check_result(res)
#> [1] TRUE

# Pretty print
print(res)

# Combine multiple results
combined <- bind_check_results(list(res_dup, res_dur, res_gps))
```

\newpage

# Function Reference by Category

This section documents every check function in the package,
organized by thematic category. For each function we provide:
the full signature with default arguments, a brief description,
a code example, and notes on the returned \texttt{check\_result}.

## Identification Checks

These checks ensure that survey records can be uniquely and
correctly identified, and that the sample frame is respected.

### `check_duplicate_ids()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_duplicate_ids(data, id_col)
```

**Description:** Identifies rows with non-unique values in
\texttt{id\_col}. All rows sharing a duplicated ID are flagged.
Severity: \texttt{error}.

**Example:**
```{r check-dup-ids}
res <- check_duplicate_ids(survey, id_col = "hh_id")
res$n_flagged
#> [1] 2
res$flagged_ids
#> [1] "HH-0003" "HH-0003"
```

**Returns:** \texttt{check\_result} with
\texttt{check\_name = "duplicate\_hh\_id"}.
\texttt{summary\_stat} includes the count of unique duplicated
values.

---

### `check_duplicate_fingerprint()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_duplicate_fingerprint(data, id_col, quasi_ids)
```

**Description:** Detects near-duplicates by comparing combinations
of quasi-identifier columns (e.g., name, date of birth, village).
Even if \texttt{id\_col} values differ, records sharing identical
quasi-identifiers are flagged. Severity: \texttt{warning}.

**Example:**
```{r check-dup-fingerprint}
res <- check_duplicate_fingerprint(
  survey, id_col = "hh_id",
  quasi_ids = c("head_name", "village", "hh_size")
)
```

**Returns:** \texttt{check\_result} with
\texttt{check\_name = "duplicate\_fingerprint"}.

---

### `check_missing_ids()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_missing_ids(data, id_col)
```

**Description:** Flags observations where \texttt{id\_col} is
\texttt{NA}, empty string, or whitespace-only. Severity:
\texttt{error}.

**Example:**
```{r check-missing-ids}
res <- check_missing_ids(survey, id_col = "hh_id")
res$n_flagged
#> [1] 0
```

---

### `check_id_format()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_id_format(data, id_col, pattern)
```

**Description:** Validates that every ID matches the supplied
regular expression \texttt{pattern}. Useful for enforcing
project-specific ID conventions (e.g., \texttt{"^HH-[0-9]\{4\}$"}).
Severity: \texttt{warning}.

**Example:**
```{r check-id-format}
res <- check_id_format(
  survey, id_col = "hh_id", pattern = "^HH-[0-9]{4}$"
)
```

---

### `check_id_in_sample()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_id_in_sample(data, id_col, sampling_frame)
```

**Description:** Cross-references collected IDs against the
pre-defined sampling frame. IDs not found in the frame are flagged
as potential replacements or unauthorized additions. Severity:
\texttt{warning}.

**Example:**
```{r check-id-sample}
frame <- data.frame(hh_id = sprintf("HH-%04d", 1:250))
res <- check_id_in_sample(survey, id_col = "hh_id", sampling_frame = frame)
```

---

## Metadata Checks

These checks validate survey-level metadata such as consent status,
form versions, completion flags, and respondent eligibility.

### `check_consent()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_consent(data, id_col, consent_col, consent_value = 1)
```

**Description:** Flags observations where the consent field does
not equal \texttt{consent\_value}. Records without consent should
typically be excluded from analysis. Severity: \texttt{error}.

**Example:**
```{r check-consent}
res <- check_consent(
  survey, id_col = "hh_id",
  consent_col = "consent", consent_value = 1
)
```

---

### `check_form_version()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_form_version(data, id_col, version_col, expected_version)
```

**Description:** Identifies submissions collected with an outdated
or incorrect form version. Severity: \texttt{warning}.

**Example:**
```{r check-form-version}
res <- check_form_version(
  survey, id_col = "hh_id",
  version_col = "form_version", expected_version = "v2.1"
)
```

---

### `check_interview_completed()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_interview_completed(data, id_col, status_col,
                          complete_value = "complete")
```

**Description:** Flags interviews that were not marked as fully
completed. Severity: \texttt{warning}.

**Example:**
```{r check-interview-completed}
res <- check_interview_completed(
  survey, id_col = "hh_id",
  status_col = "status", complete_value = "complete"
)
```

---

### `check_survey_tracking()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_survey_tracking(data, id_col, stratum_col, target_per_stratum)
```

**Description:** Compares the number of completed surveys per
stratum against the sampling targets. Reports under- and
over-sampling by stratum. Severity: \texttt{info}.

**Example:**
```{r check-tracking}
targets <- c(region_A = 50, region_B = 75, region_C = 75)
res <- check_survey_tracking(
  survey, id_col = "hh_id",
  stratum_col = "region", target_per_stratum = targets
)
```

**Returns:** \texttt{summary\_stat} includes per-stratum counts,
targets, and deviations.

---

### `check_respondent_eligibility()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_respondent_eligibility(data, id_col, age_col = NULL,
                             gender_col = NULL, min_age = 18,
                             max_age = NULL, eligible_gender = NULL)
```

**Description:** Validates that respondents meet age and/or gender
eligibility criteria. Severity: \texttt{warning}.

**Example:**
```{r check-eligibility}
res <- check_respondent_eligibility(
  survey, id_col = "hh_id",
  age_col = "resp_age", min_age = 18, max_age = 65
)
```

---

## Completeness Checks

These functions assess the extent and pattern of missing data at
the variable, enumerator, and observation levels.

### `check_missing_by_variable()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_missing_by_variable(data, id_col, threshold = 0.05)
```

**Description:** Computes the proportion of missing values for
each column. Variables exceeding \texttt{threshold} are flagged.
Severity: \texttt{warning}.

**Example:**
```{r check-missing-var}
res <- check_missing_by_variable(survey, id_col = "hh_id", threshold = 0.05)
res$summary_stat
# A list of variable names and their missingness rates
```

---

### `check_missing_by_enumerator()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_missing_by_enumerator(data, id_col, enum_col, threshold_sd = 2)
```

**Description:** For each enumerator, computes overall missingness
rates and flags those whose rate exceeds the group mean by more
than \texttt{threshold\_sd} standard deviations. Severity:
\texttt{warning}.

**Example:**
```{r check-missing-enum}
res <- check_missing_by_enumerator(
  survey, id_col = "hh_id", enum_col = "enum_id", threshold_sd = 2
)
```

---

### `check_all_missing_variables()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_all_missing_variables(data)
```

**Description:** Identifies columns where every value is
\texttt{NA}. These may indicate programming errors in the survey
instrument. Severity: \texttt{error}.

**Example:**
```{r check-all-missing}
res <- check_all_missing_variables(survey)
```

---

### `check_skip_pattern()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_skip_pattern(data, id_col, parent_col, child_col,
                   parent_value, expect_child_na = TRUE)
```

**Description:** Validates skip logic: when \texttt{parent\_col}
equals \texttt{parent\_value}, the \texttt{child\_col} should be
\texttt{NA} (if \texttt{expect\_child\_na = TRUE}) or non-missing
(if \texttt{FALSE}). Severity: \texttt{warning}.

**Example:**
```{r check-skip}
res <- check_skip_pattern(
  survey, id_col = "hh_id",
  parent_col = "has_livestock", child_col = "n_cattle",
  parent_value = 0, expect_child_na = TRUE
)
```

---

### `check_roster_completeness()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_roster_completeness(data, id_col, hh_size_col, member_count_col)
```

**Description:** Flags households where the reported household size
does not match the number of roster entries. Severity:
\texttt{warning}. Also used in logic consistency checks
(Section 5.9).

**Example:**
```{r check-roster}
res <- check_roster_completeness(
  survey, id_col = "hh_id",
  hh_size_col = "hh_size", member_count_col = "n_members"
)
```

---

## Outlier Checks

These functions detect extreme or implausible numeric values using
statistical and rule-based methods.

### `check_outliers_iqr()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_outliers_iqr(data, id_col, num_col, multiplier = 1.5)
```

**Description:** Flags values outside the interquartile range
fence: $[Q_1 - k \cdot \text{IQR},\; Q_3 + k \cdot \text{IQR}]$
where $k$ = \texttt{multiplier}. Severity: \texttt{warning}.

**Example:**
```{r check-outlier-iqr}
res <- check_outliers_iqr(
  survey, id_col = "hh_id", num_col = "income", multiplier = 1.5
)
```

**Returns:** \texttt{summary\_stat} includes \texttt{Q1},
\texttt{Q3}, \texttt{IQR}, \texttt{lower\_fence}, and
\texttt{upper\_fence}.

---

### `check_outliers_zscore()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_outliers_zscore(data, id_col, num_col, threshold = 3)
```

**Description:** Flags values with $|z| >$ \texttt{threshold},
where $z = (x - \bar{x}) / s$. Sensitive to non-normal
distributions; consider \texttt{check\_outliers\_mad()} for
skewed data. Severity: \texttt{warning}.

**Example:**
```{r check-outlier-z}
res <- check_outliers_zscore(
  survey, id_col = "hh_id", num_col = "income", threshold = 3
)
```

---

### `check_outliers_mad()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_outliers_mad(data, id_col, num_col, threshold = 3)
```

**Description:** Uses the Median Absolute Deviation (MAD) as a
robust alternative to the z-score. Flags values where
$|x - \tilde{x}| / \text{MAD} >$ \texttt{threshold}. Severity:
\texttt{warning}.

**Example:**
```{r check-outlier-mad}
res <- check_outliers_mad(
  survey, id_col = "hh_id", num_col = "income", threshold = 3
)
```

---

### `check_hard_range()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_hard_range(data, id_col, num_col, min_val, max_val)
```

**Description:** Applies hard-coded minimum and maximum bounds.
Values outside $[\texttt{min\_val}, \texttt{max\_val}]$ are
flagged. Severity: \texttt{error}.

**Example:**
```{r check-hard-range}
res <- check_hard_range(
  survey, id_col = "hh_id", num_col = "hh_size",
  min_val = 1, max_val = 30
)
```

---

### `check_negative_values()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_negative_values(data, id_col, num_col)
```

**Description:** Flags any negative values in a column that should
contain non-negative quantities (e.g., income, area, count).
Severity: \texttt{error}.

**Example:**
```{r check-negative}
res <- check_negative_values(survey, id_col = "hh_id", num_col = "income")
```

---

## Enumerator Performance Checks

These functions monitor enumerator behavior to detect low-quality
or potentially fabricated data collection patterns.

### `check_enumerator_productivity()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_enumerator_productivity(data, id_col, enum_col, date_col,
                              max_daily = 10)
```

**Description:** Counts the number of interviews per enumerator
per day. Flags enumerator-day combinations exceeding
\texttt{max\_daily}. Severity: \texttt{warning}.

**Example:**
```{r check-productivity}
res <- check_enumerator_productivity(
  survey, id_col = "hh_id", enum_col = "enum_id",
  date_col = "date", max_daily = 10
)
```

---

### `check_enumerator_duration()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_enumerator_duration(data, id_col, enum_col, duration_col,
                          threshold_sd = 2)
```

**Description:** Computes mean interview duration per enumerator
and flags those whose average deviates by more than
\texttt{threshold\_sd} standard deviations from the overall mean.
Severity: \texttt{warning}.

**Example:**
```{r check-enum-duration}
res <- check_enumerator_duration(
  survey, id_col = "hh_id", enum_col = "enum_id",
  duration_col = "duration_min", threshold_sd = 2
)
```

---

### `check_dk_rate()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_dk_rate(data, id_col, enum_col,
              dk_value = c("dk", "don't know", "-99"),
              threshold_sd = 2)
```

**Description:** Calculates the rate of "don't know" or refusal
responses per enumerator across all variables. Enumerators with
unusually high DK rates may require retraining. Severity:
\texttt{warning}.

**Example:**
```{r check-dk-rate}
res <- check_dk_rate(
  survey, id_col = "hh_id", enum_col = "enum_id",
  dk_value = c("dk", "don't know", "-99", "98"),
  threshold_sd = 2
)
```

---

### `check_enumerator_time_gap()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_enumerator_time_gap(data, id_col, enum_col, time_col,
                          min_gap_minutes = 5)
```

**Description:** Flags consecutive interviews by the same
enumerator with a time gap shorter than \texttt{min\_gap\_minutes}.
An impossibly short gap may indicate fabrication. Severity:
\texttt{warning}.

**Example:**
```{r check-time-gap}
res <- check_enumerator_time_gap(
  survey, id_col = "hh_id", enum_col = "enum_id",
  time_col = "start_time", min_gap_minutes = 5
)
```

---

### `check_straightlining()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_straightlining(data, id_col, check_cols,
                     max_identical_rate = 0.8)
```

**Description:** For each observation, computes the proportion of
identical responses across the specified columns. Observations
exceeding \texttt{max\_identical\_rate} are flagged as potential
straightliners. Severity: \texttt{warning}. Also referenced in
fabrication detection (Section 5.8).

**Example:**
```{r check-straightlining}
likert_cols <- paste0("q_", 1:15)
res <- check_straightlining(
  survey, id_col = "hh_id",
  check_cols = likert_cols, max_identical_rate = 0.8
)
```

---

## Timing Checks

These functions validate temporal aspects of data collection,
including interview duration, date validity, and collection windows.

### `check_survey_duration()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_survey_duration(data, id_col, duration_col,
                      min_duration = 15, max_duration = 120)
```

**Description:** Flags interviews whose duration falls outside
$[\texttt{min\_duration}, \texttt{max\_duration}]$ minutes.
Severity: \texttt{warning}.

**Example:**
```{r check-duration}
res <- check_survey_duration(
  survey, id_col = "hh_id", duration_col = "duration_min",
  min_duration = 15, max_duration = 120
)
```

---

### `check_collection_window()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_collection_window(data, id_col, date_col,
                        start_date, end_date)
```

**Description:** Flags interviews with dates outside the official
data collection window. Severity: \texttt{warning}.

**Example:**
```{r check-collection-window}
res <- check_collection_window(
  survey, id_col = "hh_id", date_col = "date",
  start_date = as.Date("2025-06-01"),
  end_date   = as.Date("2025-06-30")
)
```

---

### `check_future_dates()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_future_dates(data, id_col, date_col,
                   reference_date = Sys.Date())
```

**Description:** Flags dates that are in the future relative to
\texttt{reference\_date}. Severity: \texttt{error}.

**Example:**
```{r check-future-dates}
res <- check_future_dates(
  survey, id_col = "hh_id", date_col = "date"
)
```

---

### `check_survey_end_before_start()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_survey_end_before_start(data, id_col, start_col, end_col)
```

**Description:** Flags records where the end timestamp precedes
the start timestamp. Severity: \texttt{error}. Also referenced
in logic consistency checks (Section 5.9).

**Example:**
```{r check-end-before-start}
res <- check_survey_end_before_start(
  survey, id_col = "hh_id",
  start_col = "start_time", end_col = "end_time"
)
```

---

### `check_duration_by_hh_size()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_duration_by_hh_size(data, id_col, duration_col,
                          hh_size_col,
                          min_minutes_per_member = 3)
```

**Description:** Flags interviews where the duration is
implausibly low given household size. The minimum expected
duration is computed as
$\texttt{hh\_size} \times \texttt{min\_minutes\_per\_member}$.
Severity: \texttt{warning}. Also referenced in logic
consistency checks (Section 5.9).

**Example:**
```{r check-duration-hhsize}
res <- check_duration_by_hh_size(
  survey, id_col = "hh_id",
  duration_col = "duration_min", hh_size_col = "hh_size",
  min_minutes_per_member = 3
)
```

---

## GPS Validation Checks

These functions validate geographic coordinates for boundary
compliance, accuracy, duplication, and common GPS artifacts.

### `check_gps_boundary()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_gps_boundary(data, id_col, lat_col, lon_col,
                   boundary_bbox = NULL, boundary_sf = NULL)
```

**Description:** Flags GPS coordinates that fall outside the study
area boundary. The boundary can be specified as a bounding box
(named numeric vector: \texttt{xmin, ymin, xmax, ymax}) or an
\texttt{sf} polygon object. Severity: \texttt{error}.

**Example:**
```{r check-gps-boundary}
bbox <- c(xmin = -1.7, ymin = 12.1, xmax = -1.2, ymax = 12.6)
res <- check_gps_boundary(
  survey, id_col = "hh_id",
  lat_col = "lat", lon_col = "lon",
  boundary_bbox = bbox
)
```

---

### `check_gps_centroid_distance()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_gps_centroid_distance(data, id_col, lat_col, lon_col,
                            cluster_col, max_distance_km = 10)
```

**Description:** Computes the distance of each point from its
cluster centroid and flags points exceeding
\texttt{max\_distance\_km}. Severity: \texttt{warning}.

**Example:**
```{r check-gps-centroid}
res <- check_gps_centroid_distance(
  survey, id_col = "hh_id",
  lat_col = "lat", lon_col = "lon",
  cluster_col = "village", max_distance_km = 10
)
```

---

### `check_gps_accuracy()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_gps_accuracy(data, id_col, accuracy_col, max_accuracy = 50)
```

**Description:** Flags GPS readings with an accuracy value
(in meters) exceeding \texttt{max\_accuracy}. High accuracy
values indicate poor satellite reception. Severity:
\texttt{warning}.

**Example:**
```{r check-gps-accuracy}
res <- check_gps_accuracy(
  survey, id_col = "hh_id",
  accuracy_col = "gps_accuracy", max_accuracy = 50
)
```

---

### `check_gps_duplicates()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_gps_duplicates(data, id_col, lat_col, lon_col,
                     min_distance = 0.0001)
```

**Description:** Identifies pairs of records with GPS coordinates
closer than \texttt{min\_distance} degrees. Duplicate GPS points
may indicate that an enumerator did not actually travel to the
household. Severity: \texttt{warning}.

**Example:**
```{r check-gps-dup}
res <- check_gps_duplicates(
  survey, id_col = "hh_id",
  lat_col = "lat", lon_col = "lon", min_distance = 0.0001
)
```

---

### `check_gps_swap()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_gps_swap(data, id_col, lat_col, lon_col, expected_bbox)
```

**Description:** Detects cases where latitude and longitude appear
to be swapped, by checking if swapping the values brings the point
inside the \texttt{expected\_bbox}. Severity: \texttt{error}.

**Example:**
```{r check-gps-swap}
bbox <- c(xmin = -1.7, ymin = 12.1, xmax = -1.2, ymax = 12.6)
res <- check_gps_swap(
  survey, id_col = "hh_id",
  lat_col = "lat", lon_col = "lon", expected_bbox = bbox
)
```

---

### `check_gps_null_island()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_gps_null_island(data, id_col, lat_col, lon_col,
                      null_values = c(0, 999, -999))
```

**Description:** Flags records where latitude or longitude equal
sentinel null values (0, 999, $-999$, etc.), indicating missing
or default GPS readings. The coordinates (0, 0) are known as
``Null Island.'' Severity: \texttt{error}.

**Example:**
```{r check-gps-null}
res <- check_gps_null_island(
  survey, id_col = "hh_id",
  lat_col = "lat", lon_col = "lon",
  null_values = c(0, 999, -999)
)
```

---

### `check_gps_clustering()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_gps_clustering(data, id_col, lat_col, lon_col,
                     enum_col, max_same_point = 3)
```

**Description:** Flags enumerators who report more than
\texttt{max\_same\_point} interviews at identical GPS coordinates.
Severity: \texttt{warning}.

**Example:**
```{r check-gps-cluster}
res <- check_gps_clustering(
  survey, id_col = "hh_id",
  lat_col = "lat", lon_col = "lon",
  enum_col = "enum_id", max_same_point = 3
)
```

---

### `check_gps_altitude()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_gps_altitude(data, id_col, alt_col,
                   min_alt = -500, max_alt = 6000)
```

**Description:** Flags implausible altitude readings outside
$[\texttt{min\_alt}, \texttt{max\_alt}]$ meters. Severity:
\texttt{warning}.

**Example:**
```{r check-gps-alt}
res <- check_gps_altitude(
  survey, id_col = "hh_id",
  alt_col = "altitude", min_alt = 0, max_alt = 2000
)
```

---

## Fabrication Detection Checks

These advanced statistical checks are designed to detect
fabricated or manipulated survey data. They rely on mathematical
properties (Benford's Law, entropy, ICC) that genuine data
typically exhibits.

### `check_benford_first_digit()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_benford_first_digit(data, id_col, num_col)
```

**Description:** Tests the distribution of first (leading) digits
in \texttt{num\_col} against Benford's Law using a chi-squared
goodness-of-fit test. Significant deviation may suggest data
fabrication. Severity: \texttt{info}.

**Example:**
```{r check-benford1}
res <- check_benford_first_digit(
  survey, id_col = "hh_id", num_col = "income"
)
res$summary_stat$p_value
```

**Returns:** \texttt{summary\_stat} includes the chi-squared
statistic, p-value, and observed vs. expected digit frequencies.

---

### `check_benford_second_digit()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_benford_second_digit(data, id_col, num_col)
```

**Description:** Same as above, but tests the distribution of
second digits. Severity: \texttt{info}.

**Example:**
```{r check-benford2}
res <- check_benford_second_digit(
  survey, id_col = "hh_id", num_col = "income"
)
```

---

### `check_digit_preference()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_digit_preference(data, id_col, num_col)
```

**Description:** Detects heaping or rounding bias by analyzing the
distribution of terminal digits. Excessive clustering on 0 or 5
suggests that respondents or enumerators are rounding values.
Severity: \texttt{info}.

**Example:**
```{r check-digit-pref}
res <- check_digit_preference(
  survey, id_col = "hh_id", num_col = "income"
)
```

---

### `check_icc_enumerator()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_icc_enumerator(data, id_col, enum_col, num_cols,
                     max_icc = 0.15)
```

**Description:** Computes the Intra-class Correlation Coefficient
(ICC) for numeric variables, treating enumerator as the grouping
factor. A high ICC indicates that more variance is explained by
enumerator identity than by actual respondent differences, which
is a hallmark of fabrication. Severity: \texttt{warning}.

**Example:**
```{r check-icc}
res <- check_icc_enumerator(
  survey, id_col = "hh_id", enum_col = "enum_id",
  num_cols = c("income", "expenditure", "hh_size"),
  max_icc = 0.15
)
```

---

### `check_response_entropy()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_response_entropy(data, id_col, check_cols,
                       min_entropy_ratio = 0.3)
```

**Description:** Computes the normalized Shannon entropy for each
observation across \texttt{check\_cols}. Low entropy indicates
little variation in responses, suggesting straightlining or
fabrication. Severity: \texttt{warning}.

**Example:**
```{r check-entropy}
res <- check_response_entropy(
  survey, id_col = "hh_id",
  check_cols = paste0("q_", 1:20),
  min_entropy_ratio = 0.3
)
```

---

### `check_duplicate_response_patterns()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_duplicate_response_patterns(data, id_col, check_cols,
                                  similarity_threshold = 0.95)
```

**Description:** Computes pairwise similarity (proportion of
identical responses) across observations. Pairs exceeding
\texttt{similarity\_threshold} are flagged. Severity:
\texttt{warning}.

**Example:**
```{r check-dup-patterns}
res <- check_duplicate_response_patterns(
  survey, id_col = "hh_id",
  check_cols = paste0("q_", 1:20),
  similarity_threshold = 0.95
)
```

---

### `check_variance_ratio()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_variance_ratio(data, id_col, enum_col, num_cols,
                     max_ratio = 2)
```

**Description:** For each numeric variable, computes the ratio of
between-enumerator variance to within-enumerator variance.
Enumerators with ratios exceeding \texttt{max\_ratio} may be
producing artificially homogeneous data. Severity:
\texttt{warning}.

**Example:**
```{r check-var-ratio}
res <- check_variance_ratio(
  survey, id_col = "hh_id", enum_col = "enum_id",
  num_cols = c("income", "expenditure"), max_ratio = 2
)
```

---

## Logic Consistency Checks

These functions validate internal consistency between related
survey variables.

### `check_hh_composition()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_hh_composition(data, id_col, hh_size_col, roster_count_col)
```

**Description:** Flags households where the reported total
household size does not match the count of roster members.
Severity: \texttt{warning}.

**Example:**
```{r check-hh-comp}
res <- check_hh_composition(
  survey, id_col = "hh_id",
  hh_size_col = "hh_size", roster_count_col = "n_members"
)
```

---

### `check_income_expenditure()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_income_expenditure(data, id_col, income_col,
                         expenditure_col, max_ratio = 5)
```

**Description:** Flags households where the ratio of income to
expenditure (or vice versa) exceeds \texttt{max\_ratio}. Extreme
ratios may indicate data entry errors. Severity: \texttt{warning}.

**Example:**
```{r check-inc-exp}
res <- check_income_expenditure(
  survey, id_col = "hh_id",
  income_col = "income", expenditure_col = "expenditure",
  max_ratio = 5
)
```

---

### `check_age_date_consistency()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_age_date_consistency(data, id_col, age_col, dob_col,
                           tolerance_years = 1)
```

**Description:** Compares reported age against computed age from
date of birth. Discrepancies exceeding \texttt{tolerance\_years}
are flagged. Severity: \texttt{warning}.

**Example:**
```{r check-age-dob}
res <- check_age_date_consistency(
  survey, id_col = "hh_id",
  age_col = "resp_age", dob_col = "resp_dob",
  tolerance_years = 1
)
```

---

### `check_custom_logic()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_custom_logic(data, id_col, condition_expr, description)
```

**Description:** Evaluates an arbitrary R expression
(\texttt{condition\_expr}) against the data and flags rows
where the condition is \texttt{TRUE}. This is the escape hatch
for project-specific logic checks not covered by built-in
functions. Severity: \texttt{warning}.

**Example:**
```{r check-custom-logic}
res <- check_custom_logic(
  survey, id_col = "hh_id",
  condition_expr = quote(hh_size > 15 & income < 50000),
  description = "Large household with very low income"
)
```

\noindent \textit{Note:} The functions
\texttt{check\_survey\_end\_before\_start()},
\texttt{check\_duration\_by\_hh\_size()}, and
\texttt{check\_roster\_completeness()} are also classified as
logic consistency checks. Their documentation appears in Sections
5.6 and 5.3, respectively.

---

## Text Quality Checks

These functions validate open-ended text responses, names, phone
numbers, and ``other (specify)'' fields.

### `check_other_specify()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_other_specify(data, id_col, other_col)
```

**Description:** Flags ``other (specify)'' responses that are
empty, too short, or contain placeholder text (e.g., ``xxx'',
``test'', ``asdf''). Severity: \texttt{warning}.

**Example:**
```{r check-other-specify}
res <- check_other_specify(
  survey, id_col = "hh_id", other_col = "crop_other"
)
```

---

### `check_text_length()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_text_length(data, id_col, text_col,
                  min_length = 3, max_length = 500)
```

**Description:** Flags text responses shorter than
\texttt{min\_length} or longer than \texttt{max\_length}
characters. Severity: \texttt{warning}.

**Example:**
```{r check-text-length}
res <- check_text_length(
  survey, id_col = "hh_id", text_col = "comments",
  min_length = 3, max_length = 500
)
```

---

### `check_text_duplicates()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_text_duplicates(data, id_col, text_col, min_length = 10)
```

**Description:** Identifies duplicate text responses (after
trimming and case-folding) that are at least \texttt{min\_length}
characters long. Identical long-form responses across different
records may indicate copy-paste fabrication. Severity:
\texttt{warning}.

**Example:**
```{r check-text-dup}
res <- check_text_duplicates(
  survey, id_col = "hh_id", text_col = "comments",
  min_length = 10
)
```

---

### `check_name_format()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_name_format(data, id_col, name_col)
```

**Description:** Validates name fields for common issues:
all-caps, numeric characters, single characters, or known test
entries. Severity: \texttt{info}.

**Example:**
```{r check-name-format}
res <- check_name_format(
  survey, id_col = "hh_id", name_col = "head_name"
)
```

---

### `check_phone_format()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_phone_format(data, id_col, phone_col, country_code,
                   min_digits = 8, max_digits = 15)
```

**Description:** Validates phone numbers against country-specific
formatting rules: correct number of digits, valid prefix, and no
invalid characters. Severity: \texttt{warning}.

**Example:**
```{r check-phone-format}
res <- check_phone_format(
  survey, id_col = "hh_id", phone_col = "phone",
  country_code = "BF", min_digits = 8, max_digits = 8
)
```

---

## Back-Check Checks

Back-checks (re-interviews of a subsample) are a cornerstone of
survey quality assurance. These functions compare original and
back-check data.

### `check_backcheck_t1_match()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_backcheck_t1_match(data, id_col, backcheck_data, bc_id_col,
                         t1_cols, ...)
```

**Description:** Type~1 back-check: compares Type~1 variables
(e.g., gender, location) between original and back-check data.
Any mismatch on these variables signals a serious data quality
issue because they should never change. Severity: \texttt{error}.

**Example:**
```{r check-bc-t1}
res <- check_backcheck_t1_match(
  data = survey, id_col = "hh_id",
  backcheck_data = bc_data, bc_id_col = "hh_id",
  t1_cols = c("gender", "village")
)
```

---

### `check_backcheck_t2_match()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_backcheck_t2_match(data, id_col, backcheck_data, bc_id_col,
                         t2_cols, max_error_rate = 0.10,
                         enum_col = NULL, ...)
```

**Description:** Type~2 back-check: compares categorical or
yes/no responses between original and back-check. Computes
exact-match agreement rates per variable and flags enumerators
whose error rate exceeds \texttt{max\_error\_rate}. Severity:
\texttt{warning}.

**Example:**
```{r check-bc-t2}
res <- check_backcheck_t2_match(
  data = survey, id_col = "hh_id",
  backcheck_data = bc_data, bc_id_col = "hh_id",
  t2_cols = c("roof_material", "wall_material", "water_source"),
  max_error_rate = 0.10, enum_col = "enum_id"
)
```

---

### `check_backcheck_t3_match()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_backcheck_t3_match(data, id_col, backcheck_data, bc_id_col,
                         t3_cols, ...)
```

**Description:** Type~3 back-check: compares variables that are
expected to change between original and back-check (e.g.,
satisfaction, attitude). This check is purely informational and
computes change rates per variable. Severity: \texttt{info}.

**Example:**
```{r check-bc-t3}
res <- check_backcheck_t3_match(
  data = survey, id_col = "hh_id",
  backcheck_data = bc_data, bc_id_col = "hh_id",
  t3_cols = c("satisfaction", "attitude")
)
```

---

### `check_backcheck_coverage()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_backcheck_coverage(data, id_col, backcheck_data, bc_id_col,
                         strata_col = NULL, min_coverage = 0.10,
                         ...)
```

**Description:** Verifies that a sufficient percentage of original
surveys have been back-checked. If strata are provided, coverage
is computed per stratum. Severity: \texttt{warning}.

**Example:**
```{r check-bc-coverage}
res <- check_backcheck_coverage(
  data = survey, id_col = "hh_id",
  backcheck_data = bc_data, bc_id_col = "hh_id",
  strata_col = "region", min_coverage = 0.10
)
```

---

### `check_backcheck_by_enumerator()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_backcheck_by_enumerator(data, id_col, enum_col,
                              backcheck_data, bc_id_col,
                              compare_cols, threshold_sd = 2,
                              ...)
```

**Description:** Computes per-enumerator error rates across all
comparison columns and flags enumerators whose error rate is more
than \texttt{threshold\_sd} standard deviations above the team
mean. Severity: \texttt{warning}.

**Example:**
```{r check-bc-enum}
res <- check_backcheck_by_enumerator(
  data = survey, id_col = "hh_id", enum_col = "enum_id",
  backcheck_data = bc_data, bc_id_col = "hh_id",
  compare_cols = c("roof_material", "hh_size"),
  threshold_sd = 2
)
```

---

## Paradata Checks

Paradata (text audit, server logs, device metadata) provides
fine-grained quality signals at the question level.

### `check_text_audit_duration()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_text_audit_duration(text_audit, id_col, question_col,
                          duration_col, min_seconds = 1,
                          max_seconds = 300)
```

**Description:** Flags individual question responses with
implausible durations (too fast or too slow) from text audit
paradata. Severity: \texttt{warning}.

**Example:**
```{r check-ta-duration}
res <- check_text_audit_duration(
  text_audit, id_col = "hh_id",
  question_col = "question", duration_col = "duration_sec",
  min_seconds = 1, max_seconds = 300
)
```

---

### `check_text_audit_sequence()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_text_audit_sequence(text_audit, id_col, question_col,
                          expected_order)
```

**Description:** Verifies that questions were answered in the
expected order. Out-of-order responses may indicate form
manipulation. Severity: \texttt{info}.

**Example:**
```{r check-ta-sequence}
expected <- c("consent", "hh_size", "income", "expenditure")
res <- check_text_audit_sequence(
  text_audit, id_col = "hh_id",
  question_col = "question", expected_order = expected
)
```

---

### `check_speed_violations()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_speed_violations(text_audit, id_col, question_col,
                       duration_col, min_seconds_per_field = 0.5)
```

**Description:** Counts the number of fields answered faster than
\texttt{min\_seconds\_per\_field}. Interviews with many speed
violations are flagged. Severity: \texttt{warning}.

**Example:**
```{r check-speed}
res <- check_speed_violations(
  text_audit, id_col = "hh_id",
  question_col = "question", duration_col = "duration_sec",
  min_seconds_per_field = 0.5
)
```

---

### `check_field_comments()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_field_comments(data, id_col, comment_col)
```

**Description:** Reports observations that have non-empty
field-level comments, which may contain corrections, clarifications,
or flags left by enumerators. Severity: \texttt{info}.

**Example:**
```{r check-comments}
res <- check_field_comments(
  survey, id_col = "hh_id", comment_col = "enumerator_notes"
)
```

---

### `check_photo_attachment()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_photo_attachment(data, id_col, photo_col)
```

**Description:** Flags records where a required photo attachment
is missing or has an invalid path. Severity: \texttt{warning}.

**Example:**
```{r check-photo}
res <- check_photo_attachment(
  survey, id_col = "hh_id", photo_col = "dwelling_photo"
)
```

---

### `check_device_consistency()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_device_consistency(data, id_col, enum_col, device_col)
```

**Description:** Flags enumerators who used more than one device
during data collection. While not always problematic, device
switching may warrant investigation. Severity: \texttt{info}.

**Example:**
```{r check-device}
res <- check_device_consistency(
  survey, id_col = "hh_id",
  enum_col = "enum_id", device_col = "device_id"
)
```

---

## Correction Functions

These functions apply corrections to the dataset based on a
correction log and recode open-ended responses.

### `apply_corrections()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
apply_corrections(data, correction_log, id_col, variable_col,
                  old_value_col, new_value_col)
```

**Description:** Applies a correction log to the dataset. For
each row in the log, the function locates the matching record by
\texttt{id\_col} and variable name, verifies the current value
matches \texttt{old\_value\_col}, and replaces it with
\texttt{new\_value\_col}. Returns the corrected dataset along with
a log of applied and failed corrections.

**Example:**
```{r apply-corrections}
corrections <- data.frame(
  hh_id      = c("HH-0010", "HH-0015"),
  variable   = c("duration_min", "income"),
  old_value  = c("3", "1000000000"),
  new_value  = c("30", "500000"),
  stringsAsFactors = FALSE
)

corrected <- apply_corrections(
  survey, correction_log = corrections,
  id_col = "hh_id", variable_col = "variable",
  old_value_col = "old_value", new_value_col = "new_value"
)
```

---

### `check_correction_log()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
check_correction_log(data, correction_log, id_col, variable_col,
                     old_value_col)
```

**Description:** Validates a correction log before applying it.
Checks that all referenced IDs exist, variable names are valid,
and old values match current data. Returns a \texttt{check\_result}
with mismatches flagged. Severity: \texttt{error}.

**Example:**
```{r check-correction-log}
res <- check_correction_log(
  survey, correction_log = corrections,
  id_col = "hh_id", variable_col = "variable",
  old_value_col = "old_value"
)
```

---

### `recode_other_specify()`

\vspace{0.2cm}

**Signature:**
```{r, eval=FALSE}
recode_other_specify(data, id_col, other_col, recode_map,
                     target_col = NULL)
```

**Description:** Recodes free-text ``other (specify)'' responses
into existing categories using a named mapping vector. Optionally
writes recoded values into \texttt{target\_col} (the main
categorical variable). Returns the recoded dataset.

**Example:**
```{r recode-other}
recode_map <- c(
  "millet"    = "cereals",
  "sorghum"   = "cereals",
  "groundnut" = "legumes"
)

recoded <- recode_other_specify(
  survey, id_col = "hh_id",
  other_col = "crop_other", recode_map = recode_map,
  target_col = "main_crop"
)
```

\newpage

# Configuration System

The YAML configuration system allows survey managers to define all
check parameters in a single file, enabling reproducible and
auditable quality assurance.

## YAML Structure

A configuration file has three top-level sections:

```{r config-yaml, eval=FALSE}
# config.yml
project:
  name: "Household Survey 2025"
  survey_start: "2025-01-15"
  survey_end: "2025-06-30"

variables:
  id: "hh_id"
  enumerator: "enum_id"
  date: "interview_date"
  duration: "duration_min"
  gps_latitude: "gps_lat"
  gps_longitude: "gps_lon"
  gps_accuracy: "gps_accuracy"
  consent: "consent"
  form_version: "form_version"
  hh_size: "hh_size"

checks:
  identification:
    duplicate_ids: true
    missing_ids: true
    id_format:
      pattern: "^HH-[0-9]{4}$"

  metadata:
    consent:
      consent_value: 1
    form_version:
      expected_version: "v2.1"

  completeness:
    missing_by_variable:
      threshold: 0.05
    missing_by_enumerator:
      threshold_sd: 2

  outliers:
    outliers_iqr:
      columns: ["income", "expenditure", "hh_size"]
      multiplier: 1.5
    hard_range:
      - column: "hh_size"
        min_val: 1
        max_val: 30
      - column: "income"
        min_val: 0
        max_val: 50000000

  timing:
    survey_duration:
      min_duration: 15
      max_duration: 120
    collection_window:
      start_date: "2025-06-01"
      end_date: "2025-06-30"

  gps:
    gps_boundary:
      boundary_bbox: [-1.7, 12.1, -1.2, 12.6]
    gps_null_island: true
    gps_accuracy:
      max_accuracy: 50

  enumerator:
    enumerator_productivity:
      max_daily: 10
    enumerator_duration:
      threshold_sd: 2

  fabrication:
    benford_first_digit:
      columns: ["income", "expenditure"]
    icc_enumerator:
      num_cols: ["income", "expenditure", "hh_size"]
      max_icc: 0.15
```

## Configuration Functions

```{r config-functions}
# Read and validate a YAML config file
config <- read_config("config.yml")

# Validate config structure
validate_config(config)
#> Config validation passed

# Extract a nested value with a default
id_col <- cfg_get(config, "variables", "id")
max_daily <- cfg_get(config, "checks", "enumerator",
                     "productivity", "max_daily",
                     default = 10)
```

\newpage

# Orchestrator: `run_all_checks()`

The orchestrator is the primary entry point for running a complete
suite of checks in a single call.

## Signature

```{r orchestrator-sig, eval=FALSE}
run_all_checks(
  data,
  config,
  sampling_frame  = NULL,
  backcheck_data  = NULL,
  verbose         = TRUE,
  ...
)
```

## Parameters

\begin{table}[H]
\centering
\begin{tabular}{@{}lp{9cm}@{}}
\toprule
\textbf{Parameter} & \textbf{Description} \\
\midrule
\texttt{data}            & Main survey data frame \\
\texttt{config}          & Either a path to a YAML config file or an \texttt{adc\_config} object \\
\texttt{sampling\_frame} & Optional character vector of valid IDs \\
\texttt{backcheck\_data} & Optional back-check data frame \\
\texttt{verbose}         & If \texttt{TRUE}, prints progress messages \\
\texttt{...}             & Reserved for future use \\
\bottomrule
\end{tabular}
\end{table}

## Return Value

Returns an object of class \texttt{adc\_report}, which is a named
list containing:

\begin{itemize}[leftmargin=2cm]
  \item \texttt{results}: A list of \texttt{check\_result} objects
  \item \texttt{config}: The configuration used
  \item \texttt{metadata}: Timestamp, package version, data dimensions
  \item \texttt{summary}: Aggregated counts by severity and category
\end{itemize}

## Methods

```{r report-methods}
report <- run_all_checks(survey, config = config, verbose = TRUE)

# Print overview
print(report)
#> === afcestDataCheck Report ===
#> Data: 200 observations, 12 variables
#> Checks run: 18
#> Errors:   3 | Warnings: 12 | Info: 2

# Detailed summary
summary(report)
#> Category            Checks  Errors  Warnings  Info
#> identification         3       1        0       0
#> metadata               2       1        0       0
#> completeness           2       0        2       0
#> outliers               3       0        3       0
#> ...
```

\newpage

# Platform Normalization

The package supports data from five major survey platforms by
normalizing platform-specific conventions into a common format.

## Supported Platforms

\begin{table}[H]
\centering
\caption{Supported data collection platforms}
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Platform} & \textbf{GPS Format} & \textbf{Duration} & \textbf{Metadata Cols} \\
\midrule
KoBoToolbox     & String: ``lat lon alt acc''  & Seconds (integer)  & \texttt{\_submission\_time} \\
SurveyCTO       & String: ``lat lon alt acc''  & Seconds (integer)  & \texttt{starttime, endtime} \\
ODK Central     & GeoJSON or string            & Seconds            & \texttt{\_\_system/*} \\
Survey Solutions & Separate lat/lon columns    & Minutes (decimal)  & \texttt{interview\_\_*} \\
CSPro           & Separate lat/lon columns     & Varies             & \texttt{CAPI\_*} \\
\bottomrule
\end{tabular}
\end{table}

## Key Functions

```{r platform-normalize}
# Auto-detect platform and normalize
normalized <- normalize_survey_data(
  raw_data, platform = "auto"
)
# Returns data with standardized column names:
#   .id, .enum_id, .start_time, .end_time, .duration_min,
#   .lat, .lon, .gps_accuracy, .submission_date

# Parse a GPS string (KoBoToolbox / SurveyCTO format)
coords <- parse_gps_string("12.345 -1.234 300 5.2")
coords
#> $lat
#> [1] 12.345
#> $lon
#> [1] -1.234
#> $alt
#> [1] 300
#> $accuracy
#> [1] 5.2

# Detect platform from column names (internal)
# detect_platform(data) -> "kobo" | "scto" | "odk" | "suso" | "cspro"
```

\newpage

# Export Functions

## Excel Export

```{r export-excel}
export_to_excel(report, data = survey, id_col = "hh_id",
                path = "hfc_report.xlsx",
                enum_col = "enum_id", date_col = "date")
```

The Excel workbook contains five sheets:

\begin{enumerate}[leftmargin=2cm]
  \item \textbf{Dashboard} sheet: Summary statistics including total
        observations, date range, flags by category, and enumerator
        summary.
  \item \textbf{Flagged Records} sheet: Master list of all flags with
        columns for row\_number, enumerator, date, id, gps\_lat,
        gps\_lon, duration, check\_name, check\_category, variable,
        value, flag\_reason, severity, and action. Severity is
        color-coded (red = error, orange = warning, blue = info).
  \item \textbf{Per-category} sheets: One sheet per flagged category
        with full data rows for flagged IDs, merged with flag details.
  \item \textbf{Corrections} sheet: Pre-populated template with action
        dropdown (replace/drop/okay/pending) for data correction
        workflow.
  \item \textbf{Enumerator Stats} sheet: Per-enumerator performance
        metrics including surveys, flags, flag rate, errors, and
        warnings, sorted by flag rate (descending).
\end{enumerate}

## CSV Export

```{r export-csv}
export_to_csv(report, output_dir = "hfc_output/",
              data = survey, id_col = "hh_id",
              enum_col = "enum_id", date_col = "date")
```

Generates four files in \texttt{output\_dir}:

\begin{itemize}[leftmargin=2cm]
  \item \texttt{hfc\_report\_dashboard.csv}: Summary statistics with
        submissions, data quality metrics, and category breakdowns.
  \item \texttt{hfc\_report\_flagged\_records.csv}: Long-format table
        of all flagged observations with row\_number, enumerator,
        date, ID, GPS coordinates, duration, check name, reason,
        and severity.
  \item \texttt{hfc\_report\_corrections\_template.csv}: Pre-populated
        correction template with id, variable, old\_value, new\_value,
        action, and reason columns.
  \item \texttt{hfc\_report\_enumerator\_stats.csv}: Per-enumerator
        performance metrics.
\end{itemize}

\newpage

# Integration with Shiny

The standardized \texttt{check\_result} object integrates
naturally with Shiny reactive programming. Below is a minimal
example.

```{r shiny-integration}
library(shiny)
library(bslib)
library(afcestDataCheck)

ui <- page_sidebar(
  title = "HFC Dashboard",
  sidebar = sidebar(
    fileInput("upload", "Upload survey data (.csv)"),
    actionButton("run", "Run checks", class = "btn-primary")
  ),
  card(
    card_header("Results"),
    tableOutput("summary_table")
  ),
  card(
    card_header("Flagged Observations"),
    tableOutput("flagged_table")
  )
)

server <- function(input, output, session) {
  survey_data <- reactive({
    req(input$upload)
    read.csv(input$upload$datapath)
  })

  report <- eventReactive(input$run, {
    config <- read_config("config.yml")
    run_all_checks(survey_data(), config = config, verbose = FALSE)
  })

  output$summary_table <- renderTable({
    req(report())
    summary(report())
  })

  output$flagged_table <- renderTable({
    req(report())
    results <- report()$results
    flagged <- bind_check_results(results)
    flagged[flagged$n_flagged > 0, ]
  })
}

shinyApp(ui, server)
```

\newpage

# Appendix: Complete Function Index

\small

\begin{longtable}{@{}rllll@{}}
\caption{Complete function index for \textbf{afcestDataCheck} v0.3.0} \\
\toprule
\textbf{\#} & \textbf{Function} & \textbf{Category} & \textbf{Severity} & \textbf{Section} \\
\midrule
\endfirsthead
\toprule
\textbf{\#} & \textbf{Function} & \textbf{Category} & \textbf{Severity} & \textbf{Section} \\
\midrule
\endhead
\bottomrule
\endfoot

 1 & \texttt{check\_duplicate\_ids}           & Identification   & error   & 5.1 \\
 2 & \texttt{check\_duplicate\_fingerprint}    & Identification   & warning & 5.1 \\
 3 & \texttt{check\_missing\_ids}              & Identification   & error   & 5.1 \\
 4 & \texttt{check\_id\_format}                & Identification   & warning & 5.1 \\
 5 & \texttt{check\_id\_in\_sample}            & Identification   & warning & 5.1 \\
\addlinespace
 6 & \texttt{check\_consent}                   & Metadata         & error   & 5.2 \\
 7 & \texttt{check\_form\_version}             & Metadata         & warning & 5.2 \\
 8 & \texttt{check\_interview\_completed}      & Metadata         & warning & 5.2 \\
 9 & \texttt{check\_survey\_tracking}          & Metadata         & info    & 5.2 \\
10 & \texttt{check\_respondent\_eligibility}   & Metadata         & warning & 5.2 \\
\addlinespace
11 & \texttt{check\_missing\_by\_variable}     & Completeness     & warning & 5.3 \\
12 & \texttt{check\_missing\_by\_enumerator}   & Completeness     & warning & 5.3 \\
13 & \texttt{check\_all\_missing\_variables}   & Completeness     & error   & 5.3 \\
14 & \texttt{check\_skip\_pattern}             & Completeness     & warning & 5.3 \\
15 & \texttt{check\_roster\_completeness}      & Completeness     & warning & 5.3 \\
\addlinespace
16 & \texttt{check\_outliers\_iqr}             & Outliers         & warning & 5.4 \\
17 & \texttt{check\_outliers\_zscore}          & Outliers         & warning & 5.4 \\
18 & \texttt{check\_outliers\_mad}             & Outliers         & warning & 5.4 \\
19 & \texttt{check\_hard\_range}               & Outliers         & error   & 5.4 \\
20 & \texttt{check\_negative\_values}          & Outliers         & error   & 5.4 \\
\addlinespace
21 & \texttt{check\_enumerator\_productivity}  & Enumerator       & warning & 5.5 \\
22 & \texttt{check\_enumerator\_duration}      & Enumerator       & warning & 5.5 \\
23 & \texttt{check\_dk\_rate}                  & Enumerator       & warning & 5.5 \\
24 & \texttt{check\_enumerator\_time\_gap}     & Enumerator       & warning & 5.5 \\
25 & \texttt{check\_straightlining}            & Enumerator       & warning & 5.5 \\
\addlinespace
26 & \texttt{check\_survey\_duration}          & Timing           & warning & 5.6 \\
27 & \texttt{check\_collection\_window}        & Timing           & warning & 5.6 \\
28 & \texttt{check\_future\_dates}             & Timing           & error   & 5.6 \\
29 & \texttt{check\_survey\_end\_before\_start}& Timing           & error   & 5.6 \\
30 & \texttt{check\_duration\_by\_hh\_size}    & Timing           & warning & 5.6 \\
\addlinespace
31 & \texttt{check\_gps\_boundary}             & GPS              & error   & 5.7 \\
32 & \texttt{check\_gps\_centroid\_distance}   & GPS              & warning & 5.7 \\
33 & \texttt{check\_gps\_accuracy}             & GPS              & warning & 5.7 \\
34 & \texttt{check\_gps\_duplicates}           & GPS              & warning & 5.7 \\
35 & \texttt{check\_gps\_swap}                 & GPS              & error   & 5.7 \\
36 & \texttt{check\_gps\_null\_island}         & GPS              & error   & 5.7 \\
37 & \texttt{check\_gps\_clustering}           & GPS              & warning & 5.7 \\
38 & \texttt{check\_gps\_altitude}             & GPS              & warning & 5.7 \\
\addlinespace
39 & \texttt{check\_benford\_first\_digit}     & Fabrication      & info    & 5.8 \\
40 & \texttt{check\_benford\_second\_digit}    & Fabrication      & info    & 5.8 \\
41 & \texttt{check\_digit\_preference}         & Fabrication      & info    & 5.8 \\
42 & \texttt{check\_icc\_enumerator}           & Fabrication      & warning & 5.8 \\
43 & \texttt{check\_response\_entropy}         & Fabrication      & warning & 5.8 \\
44 & \texttt{check\_duplicate\_response\_patterns} & Fabrication  & warning & 5.8 \\
45 & \texttt{check\_variance\_ratio}           & Fabrication      & warning & 5.8 \\
\addlinespace
46 & \texttt{check\_hh\_composition}           & Logic            & warning & 5.9 \\
47 & \texttt{check\_income\_expenditure}       & Logic            & warning & 5.9 \\
48 & \texttt{check\_age\_date\_consistency}    & Logic            & warning & 5.9 \\
49 & \texttt{check\_custom\_logic}             & Logic            & warning & 5.9 \\
\addlinespace
50 & \texttt{check\_other\_specify}            & Text             & warning & 5.10 \\
51 & \texttt{check\_text\_length}              & Text             & warning & 5.10 \\
52 & \texttt{check\_text\_duplicates}          & Text             & warning & 5.10 \\
53 & \texttt{check\_name\_format}              & Text             & info    & 5.10 \\
54 & \texttt{check\_phone\_format}             & Text             & warning & 5.10 \\
\addlinespace
55 & \texttt{check\_backcheck\_t1\_match}       & Back-Check       & error   & 5.11 \\
56 & \texttt{check\_backcheck\_t2\_match}       & Back-Check       & warning & 5.11 \\
57 & \texttt{check\_backcheck\_t3\_match}       & Back-Check       & info    & 5.11 \\
58 & \texttt{check\_backcheck\_coverage}       & Back-Check       & warning & 5.11 \\
59 & \texttt{check\_backcheck\_by\_enumerator} & Back-Check       & warning & 5.11 \\
\addlinespace
60 & \texttt{check\_text\_audit\_duration}     & Paradata         & warning & 5.12 \\
61 & \texttt{check\_text\_audit\_sequence}     & Paradata         & info    & 5.12 \\
62 & \texttt{check\_speed\_violations}         & Paradata         & warning & 5.12 \\
63 & \texttt{check\_field\_comments}           & Paradata         & info    & 5.12 \\
64 & \texttt{check\_photo\_attachment}         & Paradata         & warning & 5.12 \\
65 & \texttt{check\_device\_consistency}       & Paradata         & info    & 5.12 \\
\addlinespace
66 & \texttt{apply\_corrections}               & Corrections      & ---     & 5.13 \\
67 & \texttt{check\_correction\_log}           & Corrections      & error   & 5.13 \\
68 & \texttt{recode\_other\_specify}           & Corrections      & ---     & 5.13 \\
\addlinespace
69 & \texttt{new\_check\_result}               & Infrastructure   & ---     & 4 \\
70 & \texttt{is\_check\_result}                & Infrastructure   & ---     & 4 \\
71 & \texttt{bind\_check\_results}             & Infrastructure   & ---     & 4 \\
72 & \texttt{read\_config}                     & Configuration    & ---     & 6 \\
73 & \texttt{validate\_config}                 & Configuration    & ---     & 6 \\
74 & \texttt{cfg\_get}                         & Configuration    & ---     & 6 \\
75 & \texttt{run\_all\_checks}                 & Orchestrator     & ---     & 8 \\
76 & \texttt{normalize\_survey\_data}          & Platform         & ---     & 9 \\
77 & \texttt{parse\_gps\_string}               & Platform         & ---     & 9 \\
78 & \texttt{export\_to\_excel}                & Export           & ---     & 10 \\
79 & \texttt{export\_to\_csv}                  & Export           & ---     & 10 \\

\end{longtable}

\normalsize

\vspace{1cm}

\begin{center}
\rule{0.5\textwidth}{0.4pt}\\[0.5cm]
\textit{afcestDataCheck} is developed and maintained by\\
\textbf{AfCEST --- African Center for Studies and Training}\\
Ouagadougou, Burkina Faso\\[0.3cm]
\texttt{https://github.com/afcest/afcestDataCheck}
\end{center}
