---
title: "Getting Started with afcestDataCheck"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with afcestDataCheck}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## 1. Introduction

**afcestDataCheck** is an R package for automated high-frequency data quality
checks on survey data. It is designed for field coordinators, data managers,
and M&E specialists working on agricultural, health, and socioeconomic surveys
in development projects.

### Why another HFC package?

Several excellent tools exist for high-frequency checks:

| Tool | Language | Checks | Fabrication detection | Multi-platform |
|------|----------|--------|-----------------------|----------------|
| `iehfc` (World Bank) | Stata | ~15 | No | Stata-only |
| `ipacheck` (IPA) | Stata | ~20 | Limited | Stata-only |
| `HighFrequencyChecks` (UNHCR) | R | ~25 | No | KoBo-focused |
| **afcestDataCheck** | **R** | **54+** | **Yes (7 methods)** | **5 platforms** |

### What makes afcestDataCheck different

- **54+ check functions** organized into 10 categories
- **Fabrication detection**: Benford's Law (1st and 2nd digit), digit preference,
  intraclass correlation (ICC), variance ratio, response entropy, duplicate
  response patterns, and straightlining
- **Multi-platform normalizer**: KoBoToolbox, SurveyCTO, ODK Central,
  Survey Solutions, and CSPro -- auto-detected or manually specified
- **YAML-driven configuration**: one config file controls which checks run,
  what thresholds apply, and which variables to target
- **Standardized output**: every check returns the same `check_result` structure,
  making results easy to combine, filter, and export
- **Production-ready exports**: styled Excel workbooks with conditional formatting,
  and CSV outputs for automated pipelines

## 2. Installation

```{r install}
# Install from GitHub (requires the remotes package)
install.packages("remotes")
remotes::install_github("afcest/survey-data-quality")

# For Excel export support, also install openxlsx2
install.packages("openxlsx2")
```

Load the package:

```{r library}
library(afcestDataCheck)
```

## 3. Quick Start (5-minute workflow)

This section walks you through a complete workflow using synthetic data.

### 3.1 Create sample data

```{r sample-data}
library(afcestDataCheck)

survey_data <- data.frame(
  hh_id = c("HH001", "HH002", "HH003", "HH003", "HH004",
             NA, "HH006", "HH007", "HH008", "HH009"),
  enum_id = c("E01", "E01", "E02", "E02", "E01",
               "E02", "E03", "E03", "E03", "E03"),
  submission_date = as.Date(c(
    "2025-02-01", "2025-02-01", "2025-02-02", "2025-02-02", "2025-02-03",
    "2025-02-03", "2025-02-04", "2025-02-04", "2025-02-04", "2025-02-04"
  )),
  duration_minutes = c(25, 5, 45, 30, 120, 15, 22, 28, 35, 40),
  income = c(50000, 45000, 60000, -5000, 1000000,
             30000, 42000, 55000, 48000, 51000),
  expenditure = c(40000, 20000, 50000, 80000, 60000,
                  25000, 38000, 52000, 44000, 49000),
  hh_size = c(5, 3, 7, 4, 2, 50, 6, 4, 8, 3),
  gps_latitude = c(12.37, 12.38, 12.36, 12.36, 0,
                    12.40, 12.39, 12.39, 12.41, 12.42),
  gps_longitude = c(-1.52, -1.53, -1.51, -1.51, 0,
                     -1.55, -1.54, -1.54, -1.56, -1.57),
  respondent_name = c("Adama Ouedraogo", "Fatima Diallo", "test", "Ibrahim Sow",
                       "A", "Mariam Traore", "Ousmane Belem", "Aissata Kone",
                       "Boureima Sanou", "Salamata Zoungrana"),
  phone = c("70123456", "76543210", "123", "71234567", "70000000",
            "76000000", "78123456", "70111111", "76222222", "78333333"),
  stringsAsFactors = FALSE
)
```

### 3.2 Run individual checks

Each check function takes a data frame, an ID column name, and check-specific
parameters. Every function returns a `check_result` object:

```{r individual-checks}
# Check for duplicate IDs
res_dup <- check_duplicate_ids(survey_data, "hh_id")
print(res_dup)
#> [x] A01_duplicate_id: 1/10 flagged (error)

# Check for missing IDs
res_miss <- check_missing_ids(survey_data, "hh_id")
print(res_miss)
#> [x] A03_missing_id: 1/10 flagged (error)

# Check survey duration
res_dur <- check_survey_duration(
  survey_data, "hh_id", "duration_minutes",
  min_duration = 15, max_duration = 90
)
print(res_dur)

# Check outliers using IQR
res_out <- check_outliers_iqr(survey_data, "hh_id", "income", multiplier = 1.5)
print(res_out)

# Access flagged IDs programmatically
res_dup$flagged_ids
#> [1] "HH003"

res_dup$flag_reason
#> [1] "ID 'HH003' appears 2 times"
```

### 3.3 Run all checks from a config file

The most powerful way to use the package is through the orchestrator,
`run_all_checks()`, which reads a YAML config file and runs every
enabled check:

```{r run-all}
report <- run_all_checks(survey_data, "path/to/config.yml")

# Print a summary
print(report)
#> -- afcestDataCheck Report: Example Household Survey --
#> Timestamp: 2025-02-04 14:30:00
#> Checks run: 12
#> Checks with flags: 8
#> Check errors: 0

# Get the summary tibble
summary(report)
#> # A tibble: 12 x 6
#>    check_name            check_category n_flagged n_total pct_flagged severity
#>    <chr>                 <chr>              <int>   <int>       <dbl> <chr>
#>  1 A01_duplicate_id      identification         1      10       10    error
#>  2 A03_missing_id        identification         1      10       10    error
#>  ...
```

### 3.4 Export results

```{r export}
# Export to a styled Excel workbook
export_to_excel(
  report$results,
  survey_data,
  id_col = "hh_id",
  path = "hfc_report.xlsx"
)

# Export to CSV files
export_to_csv(
  report$results,
  output_dir = "output/",
  prefix = "hfc_report"
)
# Creates: output/hfc_report_summary.csv
#          output/hfc_report_flagged.csv
```

## 4. Configuration

### 4.1 The YAML config structure

afcestDataCheck uses a YAML file with three required sections and several
optional ones:

```yaml
# ---- Required sections ----
project:
  name: "Household Survey 2025"
  survey_start: "2025-01-15"
  survey_end: "2025-03-15"

variables:
  id: "hh_id"                  # Required
  enumerator: "enumerator_id"  # Required
  date: "submission_date"
  start_time: "start"
  end_time: "end"
  duration: "duration_minutes"

# ---- Optional: check configuration ----
checks:
  identification:
    duplicate_ids: true
    missing_ids: true
  completeness:
    missing_by_variable:
      enabled: true
      threshold: 0.05
  outliers:
    method: "iqr"
    multiplier: 1.5
    variables:
      - col: "income"
        min: 0
        max: 1000000
  timing:
    duration:
      min_minutes: 15
      max_minutes: 120
```

### 4.2 Minimal vs. full config

A **minimal config** needs only the `project` and `variables` sections.
All checks with `default_val = TRUE` in the orchestrator will run
automatically (duplicates, missing IDs, missing by variable, duration, etc.):

```yaml
# minimal_config.yml
project:
  name: "Quick check"
variables:
  id: "hh_id"
  enumerator: "enum_id"
```

A **full config** explicitly enables/disables every check and sets all
thresholds. See `inst/example_config.yml` for the complete template.

### 4.3 Enabling and disabling checks

Checks can be toggled using a simple boolean or an `enabled` flag inside
a settings block:

```yaml
checks:
  identification:
    duplicate_ids: true        # Simple boolean
    id_in_sample: false        # Disabled
    id_format:
      enabled: true            # Nested enabled flag
      pattern: "^HH\\d{3}$"   # Check-specific parameter

  enumerator:
    productivity:              # This entire block enables the check
      max_daily: 8             # and sets its threshold
```

### 4.4 Setting thresholds

Thresholds are set per-check inside the YAML:

```yaml
checks:
  outliers:
    method: "iqr"       # Choose: "iqr", "zscore", or "mad"
    multiplier: 1.5     # IQR fence multiplier (used with "iqr")
    threshold: 3        # Z-score / MAD threshold (used with "zscore" or "mad")

  timing:
    duration:
      min_minutes: 15   # Surveys shorter than this are flagged
      max_minutes: 120  # Surveys longer than this are flagged

  enumerator:
    productivity:
      max_daily: 10     # Max surveys per enumerator per day
    dk_rate:
      dk_values: ["dk", "don't know", "-99", "-88"]
      threshold_sd: 2   # Flag if DK rate > mean + 2*SD
```

### 4.5 Reading config in code

```{r config-api}
# Read and validate
cfg <- read_config("config.yml")

# Access nested values safely with defaults
method <- cfg_get(cfg, "checks", "outliers", "method", default = "iqr")
max_daily <- cfg_get(cfg, "checks", "enumerator", "productivity",
                     "max_daily", default = 10)
```

## 5. Check Categories

afcestDataCheck organizes its 54+ checks into 10 categories. Below are
examples for each.

### 5.1 Identification

Detect duplicate, missing, or invalid survey IDs.

```{r cat-identification}
# A.01: Duplicate primary keys
check_duplicate_ids(survey_data, "hh_id")

# A.03: Missing or empty IDs
check_missing_ids(survey_data, "hh_id")

# A.02: Fingerprint duplicates (different ID, same quasi-identifiers)
check_duplicate_fingerprint(
  survey_data, "hh_id",
  quasi_ids = c("gps_latitude", "gps_longitude", "respondent_name")
)

# A.05: IDs not in the sampling frame
valid_ids <- c("HH001", "HH002", "HH003", "HH004", "HH006")
check_id_in_sample(survey_data, "hh_id", sampling_frame = valid_ids)
```

### 5.2 Metadata

Validate consent, form version, interview status, and survey tracking.

```{r cat-metadata}
# A.06: Consent check
check_consent(survey_data, "hh_id", "consent", consent_value = 1)

# A.07: Form version validation
check_form_version(survey_data, "hh_id", "form_version",
                   expected_version = "v2.1")

# A.09: Interview completion status
check_interview_completed(survey_data, "hh_id", "status",
                          complete_value = "complete")

# A.10: Survey tracking by stratum
check_survey_tracking(survey_data, "hh_id", "region",
                      target_per_stratum = list(North = 100, South = 150))
```

### 5.3 Completeness

Monitor missing data rates and validate skip patterns.

```{r cat-completeness}
# C.01: Missing rate by variable (flag variables with >5% missing)
check_missing_by_variable(survey_data, "hh_id", threshold = 0.05)

# C.02: Missing rate by enumerator (flag enumerators 2 SD above mean)
check_missing_by_enumerator(survey_data, "hh_id", "enum_id",
                            threshold_sd = 2)

# C.04: Entirely missing variables
check_all_missing_variables(survey_data)

# C.03: Skip pattern validation
check_skip_pattern(
  survey_data, "hh_id",
  parent_col = "has_livestock",
  child_col = "n_cattle",
  parent_value = "no",
  expect_child_na = TRUE
)

# C.07: Roster completeness (HH size vs. roster member count)
check_roster_completeness(survey_data, "hh_id",
                          hh_size_col = "hh_size",
                          member_count_col = "roster_count")
```

### 5.4 Outliers

Detect statistical and hard-range outliers.

```{r cat-outliers}
# D.01: IQR method (flag values outside Q1 - 1.5*IQR, Q3 + 1.5*IQR)
check_outliers_iqr(survey_data, "hh_id", "income", multiplier = 1.5)

# D.02: Robust Z-score (median/MAD-based)
check_outliers_zscore(survey_data, "hh_id", "income", threshold = 3)

# D.03: Modified Z-score (MAD method)
check_outliers_mad(survey_data, "hh_id", "income", threshold = 3)

# D.04: Hard range constraints
check_hard_range(survey_data, "hh_id", "hh_size", min_val = 1, max_val = 30)

# D.09: Negative values in non-negative columns
check_negative_values(survey_data, "hh_id", "income")
```

### 5.5 Enumerator Performance

Monitor enumerator behavior for quality and productivity issues.

```{r cat-enumerator}
# B.01: Daily productivity (flag if > max_daily surveys per day)
check_enumerator_productivity(survey_data, "hh_id", "enum_id",
                              "submission_date", max_daily = 10)

# B.02: Mean duration by enumerator (flag if suspiciously low)
check_enumerator_duration(survey_data, "hh_id", "enum_id",
                          "duration_minutes", threshold_sd = 2)

# B.03: "Don't know" response rate by enumerator
check_dk_rate(survey_data, "hh_id", "enum_id",
              dk_value = c("dk", "don't know", "-99"),
              threshold_sd = 2)

# B.08: Time gap between consecutive surveys
check_enumerator_time_gap(survey_data, "hh_id", "enum_id",
                          "start_time", min_gap_minutes = 5)
```

### 5.6 Timing

Validate survey duration and collection dates.

```{r cat-timing}
# F.01: Survey duration bounds
check_survey_duration(survey_data, "hh_id", "duration_minutes",
                      min_duration = 15, max_duration = 120)

# F.02: Collection window (outside survey period)
check_collection_window(survey_data, "hh_id", "submission_date",
                        start_date = "2025-01-15", end_date = "2025-03-15")

# F.03: Future dates
check_future_dates(survey_data, "hh_id", "submission_date")

# F.04: End time before start time
check_survey_end_before_start(survey_data, "hh_id", "start", "end")

# F.06: Duration relative to household size
check_duration_by_hh_size(survey_data, "hh_id", "duration_minutes",
                          "hh_size", min_minutes_per_member = 3)
```

### 5.7 GPS Validation

Validate geographic coordinates for accuracy, boundaries, and anomalies.

```{r cat-gps}
# E.01: GPS boundary (bounding box or sf polygon)
check_gps_boundary(survey_data, "hh_id", "gps_latitude", "gps_longitude",
                   boundary_bbox = list(xmin = -6, xmax = 3,
                                        ymin = 9, ymax = 15))

# E.02: Distance from cluster centroid
check_gps_centroid_distance(survey_data, "hh_id",
                            "gps_latitude", "gps_longitude",
                            cluster_col = "cluster_id",
                            max_distance_km = 10)

# E.03: GPS accuracy threshold
check_gps_accuracy(survey_data, "hh_id", "gps_accuracy",
                   max_accuracy = 50)

# E.04: Duplicate GPS coordinates
check_gps_duplicates(survey_data, "hh_id",
                     "gps_latitude", "gps_longitude",
                     min_distance = 0.0001)

# E.05: Swapped latitude/longitude
check_gps_swap(survey_data, "hh_id", "gps_latitude", "gps_longitude",
               expected_bbox = list(lat_min = 9, lat_max = 15,
                                    lon_min = -6, lon_max = 3))

# E.07: Null island / default GPS values (0, 999, -999)
check_gps_null_island(survey_data, "hh_id",
                      "gps_latitude", "gps_longitude")

# E.08: Excessive GPS clustering (fabrication indicator)
check_gps_clustering(survey_data, "hh_id",
                     "gps_latitude", "gps_longitude",
                     enum_col = "enum_id", max_same_point = 3)

# E.10: Altitude plausibility
check_gps_altitude(survey_data, "hh_id", "gps_altitude",
                   min_alt = -500, max_alt = 6000)
```

### 5.8 Fabrication Detection

Statistical tests to identify potentially fabricated survey data.

```{r cat-fabrication}
# M.01: Benford's Law first-digit test
check_benford_first_digit(survey_data, "hh_id", "income")

# M.02: Benford's Law second-digit test
check_benford_second_digit(survey_data, "hh_id", "income")

# M.03: Terminal digit preference (digit heaping)
check_digit_preference(survey_data, "hh_id", "income")

# B.09: Straightlining (identical responses across columns)
check_straightlining(
  survey_data, "hh_id",
  check_cols = c("q1", "q2", "q3", "q4", "q5"),
  max_identical_rate = 0.8
)

# M.04: Intraclass correlation by enumerator
check_icc_enumerator(
  survey_data, "hh_id", "enum_id",
  num_cols = c("income", "expenditure", "hh_size"),
  max_icc = 0.15
)

# M.06: Response entropy (low diversity detection)
check_response_entropy(
  survey_data, "hh_id",
  check_cols = c("q1", "q2", "q3", "q4", "q5"),
  min_entropy_ratio = 0.3
)

# M.07: Duplicate response patterns across surveys
check_duplicate_response_patterns(
  survey_data, "hh_id",
  check_cols = c("income", "expenditure", "hh_size"),
  similarity_threshold = 0.95
)

# M.08: Between/within variance ratio by enumerator
check_variance_ratio(
  survey_data, "hh_id", "enum_id",
  num_cols = c("income", "expenditure"),
  max_ratio = 2
)
```

### 5.9 Logic Consistency

Cross-variable logical consistency checks.

```{r cat-logic}
# G.01: Household composition plausibility
check_hh_composition(survey_data, "hh_id", "hh_size",
                     roster_count_col = "roster_count")

# G.04: Income-expenditure ratio
check_income_expenditure(survey_data, "hh_id",
                         "income", "expenditure",
                         max_ratio = 5)

# F.09: Age vs. date of birth consistency
check_age_date_consistency(survey_data, "hh_id",
                           "age", "date_of_birth",
                           tolerance_years = 1)

# G.12: Custom logic assertions (write any R expression)
check_custom_logic(
  survey_data, "hh_id",
  condition_expr = "hh_size > 0 & income < 0",
  description = "Positive HH size but negative income"
)

# Another custom check: expenditure > income for low-income households
check_custom_logic(
  survey_data, "hh_id",
  condition_expr = "income < 20000 & expenditure > 3 * income",
  description = "Very high expenditure relative to low income"
)
```

### 5.10 Text Quality

Validate free-text fields, names, and phone numbers.

```{r cat-text}
# O.01: Other (specify) write-in responses for review
check_other_specify(survey_data, "hh_id", "crop_other")

# O.02: Text response length
check_text_length(survey_data, "hh_id", "comments",
                  min_length = 3, max_length = 500)

# O.03: Duplicated text responses (copy-paste detection)
check_text_duplicates(survey_data, "hh_id", "comments",
                      min_length = 10)

# O.04: Name format validation (catches "test", digits-only, single chars)
check_name_format(survey_data, "hh_id", "respondent_name")

# O.05: Phone number format
check_phone_format(survey_data, "hh_id", "phone",
                   country_code = "+226",
                   min_digits = 8, max_digits = 15)
```

## 6. Working with Survey Platforms

### 6.1 The normalizer

`normalize_survey_data()` detects the survey platform from column names
and creates standardized columns (`.id`, `.enum`, `.date`, `.start`, `.end`,
`.duration`, `.latitude`, `.longitude`):

```{r normalize}
# Auto-detect platform
normalized <- normalize_survey_data(kobo_data, platform = "auto")

# Explicit platform
normalized <- normalize_survey_data(surveycto_data, platform = "surveycto")

# Manual column mapping (overrides platform defaults)
normalized <- normalize_survey_data(
  my_data,
  platform = "auto",
  id_col = "household_key",
  enum_col = "interviewer",
  date_col = "survey_date"
)
```

### 6.2 Supported platforms

| Platform | Auto-detection keys | ID default | Date default |
|----------|---------------------|------------|--------------|
| KoBoToolbox | `_id`, `_submission_time` | `_id` | `_submission_time` |
| SurveyCTO | `KEY`, `submissiondate` | `KEY` | `submissiondate` |
| ODK Central | `__id`, `__system/*` | `__id` | `__system/submissionDate` |
| Survey Solutions | `interview__id` | `interview__id` | `date` |
| CSPro | `guid`, `case_id` | `guid` or `case_id` | -- |

### 6.3 KoBoToolbox workflow

```{r kobo-workflow}
# If using robotoolbox to pull data directly:
# library(robotoolbox)
# kobo_data <- kobo_submissions("asset_uid")

# Or read exported CSV
kobo_data <- read.csv("kobo_export.csv")

# Normalize (auto-detects KoBo from _id and _submission_time columns)
normalized <- normalize_survey_data(kobo_data)

# Run checks using normalized column names in config
report <- run_all_checks(normalized, "config_kobo.yml")
```

### 6.4 SurveyCTO workflow

```{r surveycto-workflow}
# Read SurveyCTO export
scto_data <- read.csv("surveycto_export.csv")

# Normalize (auto-detects from KEY and submissiondate columns)
normalized <- normalize_survey_data(scto_data, platform = "surveycto")

# Your config references SurveyCTO column names
report <- run_all_checks(normalized, "config_scto.yml")
```

### 6.5 ODK Central workflow

```{r odk-workflow}
# If using ruODK:
# library(ruODK)
# odk_data <- odata_submission_get(fid = "my_form")

odk_data <- read.csv("odk_export.csv")
normalized <- normalize_survey_data(odk_data, platform = "odk")
report <- run_all_checks(normalized, "config_odk.yml")
```

### 6.6 GPS string parsing

KoBo and ODK sometimes store GPS as a single space-separated string
(`"12.37 -1.52 300 5"`). Use `parse_gps_string()` to extract components:

```{r gps-parse}
gps_strings <- c("12.37 -1.52 300 5", "12.38 -1.53 310 8", NA)
parsed <- parse_gps_string(gps_strings)
parsed
#> # A tibble: 3 x 4
#>   latitude longitude altitude accuracy
#>      <dbl>     <dbl>    <dbl>    <dbl>
#> 1    12.37     -1.52      300        5
#> 2    12.38     -1.53      310        8
#> 3    NA        NA          NA       NA
```

## 7. The `check_result` Object

### 7.1 Structure

Every check function returns a `check_result` object -- a named list with
a consistent structure:

```{r result-structure}
res <- check_duplicate_ids(survey_data, "hh_id")

str(res)
#> List of 9
#>  $ check_name    : chr "A01_duplicate_id"
#>  $ check_category: chr "identification"
#>  $ n_flagged     : int 1
#>  $ n_total       : int 10
#>  $ flagged_ids   : chr "HH003"
#>  $ flag_reason   : chr "ID 'HH003' appears 2 times"
#>  $ severity      : chr "error"
#>  $ summary_stat  : list (check-specific statistics)
#>  $ timestamp     : POSIXct
#>  - attr(*, "class")= chr "check_result"
```

Key fields:

- **`flagged_ids`**: character vector of IDs that failed the check
- **`flag_reason`**: human-readable reason per flagged ID
- **`severity`**: `"error"` (must fix), `"warning"` (investigate), `"info"` (review)
- **`summary_stat`**: a list of check-specific statistics (bounds, means, counts)

### 7.2 Accessing flagged IDs

```{r access-flagged}
# Get flagged IDs
res$flagged_ids
#> [1] "HH003"

# Get reasons
res$flag_reason
#> [1] "ID 'HH003' appears 2 times"

# Filter original data to flagged records
flagged_rows <- survey_data[survey_data$hh_id %in% res$flagged_ids, ]

# Check severity
res$severity
#> [1] "error"

# Access summary statistics
res$summary_stat$n_duplicated_groups
#> [1] 1
```

### 7.3 Combining results with `bind_check_results()`

Combine multiple `check_result` objects into a single summary tibble:

```{r bind-results}
res1 <- check_duplicate_ids(survey_data, "hh_id")
res2 <- check_missing_ids(survey_data, "hh_id")
res3 <- check_survey_duration(survey_data, "hh_id", "duration_minutes",
                              min_duration = 15, max_duration = 90)

summary_tbl <- bind_check_results(res1, res2, res3)
summary_tbl
#> # A tibble: 3 x 6
#>   check_name         check_category n_flagged n_total pct_flagged severity
#>   <chr>              <chr>              <int>   <int>       <dbl> <chr>
#> 1 A01_duplicate_id   identification         1      10       10    error
#> 2 A03_missing_id     identification         1      10       10    error
#> 3 F01_survey_duration timing                 2      10       20    warning

# Also accepts a list of results
all_results <- list(res1, res2, res3)
summary_tbl <- bind_check_results(all_results)
```

### 7.4 Checking and testing

```{r is-check}
is_check_result(res1)
#> [1] TRUE

is_check_result(data.frame())
#> [1] FALSE
```

## 8. Exporting Results

### 8.1 Excel export (for field supervisors)

`export_to_excel()` creates a professional, multi-sheet workbook:

- **Summary** sheet: one row per check, with conditional formatting (red for
  errors, yellow for warnings)
- **Flagged Records** sheet: long-format table of every flagged ID with
  the check name, reason, and severity
- **Per-category sheets**: the original data rows for flagged IDs, grouped
  by check category (Identification, Timing, GPS, etc.)

```{r export-excel}
report <- run_all_checks(survey_data, "config.yml")

export_to_excel(
  report = report$results,
  data = survey_data,
  id_col = "hh_id",
  path = "hfc_report_2025-02-04.xlsx"
)
#> v Excel report saved to hfc_report_2025-02-04.xlsx
```

The workbook is ready to share with field supervisors via email or WhatsApp.

### 8.2 CSV export (for data managers and pipelines)

`export_to_csv()` writes two files -- a summary and a flagged-records table --
suitable for automated pipelines, version control, or database ingestion:

```{r export-csv}
export_to_csv(
  report = report$results,
  output_dir = "output/daily_checks/",
  prefix = "hfc_2025-02-04"
)
#> v Summary written to output/daily_checks/hfc_2025-02-04_summary.csv
#> v Flagged records written to output/daily_checks/hfc_2025-02-04_flagged.csv
```

## 9. Back-Check Workflow

Back-checks (a.k.a. spot-checks or re-interviews) involve resurveying a
subset of households to verify original data. afcestDataCheck supports
three tiers:

| Tier | Purpose | Method |
|------|---------|--------|
| T1 | Verify enumerator visited the household | GPS distance comparison |
| T2 | Verify key categorical responses match | Exact match on select questions |
| T3 | Verify key continuous responses are close | Difference within tolerance |

```{r backcheck}
# Assume you have original and back-check data
original <- read.csv("data/main_survey.csv")
backcheck <- read.csv("data/backcheck_survey.csv")

# The orchestrator can accept backcheck_data as a parameter
report <- run_all_checks(
  original,
  config = "config.yml",
  backcheck_data = backcheck
)
```

Back-check configuration is done in the YAML under the `backcheck` section.
This lets you specify which columns to compare and the acceptable tolerance
for continuous variables.

## 10. Advanced: Custom Logic Checks

### 10.1 Using `check_custom_logic()` with expressions

The `check_custom_logic()` function accepts any valid R expression as a
string. The expression is evaluated in the context of the data frame, so
column names are available directly. Rows where the expression evaluates to
`TRUE` are flagged as violations:

```{r custom-logic}
# Flag households where number of children exceeds household size
check_custom_logic(
  survey_data, "hh_id",
  condition_expr = "n_children > hh_size",
  description = "Number of children exceeds total household size"
)

# Flag farmers with zero land area but non-zero crop production
check_custom_logic(
  survey_data, "hh_id",
  condition_expr = "is_farmer == 1 & plot_area_ha == 0 & crop_production_kg > 0",
  description = "Farmer reports zero land but positive crop production"
)

# Flag implausible yield (kg/ha)
check_custom_logic(
  survey_data, "hh_id",
  condition_expr = "plot_area_ha > 0 & (crop_production_kg / plot_area_ha) > 10000",
  description = "Crop yield exceeds 10,000 kg/ha (implausible)"
)
```

### 10.2 Using custom checks in the config

You can define custom logic checks directly in your YAML config, and the
orchestrator will run them automatically:

```yaml
checks:
  logic:
    custom:
      - condition: "n_children > hh_size"
        description: "Children exceed household size"
      - condition: "is_farmer == 1 & plot_area_ha == 0 & crop_production_kg > 0"
        description: "Farmer with zero land but positive production"
      - condition: "income < 0 & hh_size > 0"
        description: "Negative income for valid household"
```

### 10.3 Skip pattern validation

Use `check_skip_pattern()` to validate conditional logic in your
questionnaire. For example, if a respondent says "no" to owning livestock,
the livestock count should be NA:

```{r skip-pattern}
check_skip_pattern(
  survey_data, "hh_id",
  parent_col = "has_livestock",
  child_col = "n_cattle",
  parent_value = "no",
  expect_child_na = TRUE
)
# Flags records where has_livestock == "no" but n_cattle is NOT NA

# You can also define skip patterns in the config YAML:
```

```yaml
skip_patterns:
  - parent_col: "has_livestock"
    child_col: "n_cattle"
    parent_value: "no"
    expect_child_na: true
  - parent_col: "is_farmer"
    child_col: "crop_type"
    parent_value: 0
    expect_child_na: true
```

## 11. Integration with Shiny Dashboards

afcestDataCheck is the data engine behind the AfCEST HFC dashboards. The
`check_result` and `adc_report` objects are designed to work seamlessly
in reactive Shiny contexts.

### 11.1 Using results in a Shiny app

```{r shiny-integration}
library(shiny)
library(bslib)
library(afcestDataCheck)

server <- function(input, output, session) {

  # Reactive: run checks when data or config changes
  report <- reactive({
    req(input$data_file)
    data <- read.csv(input$data_file$datapath)
    run_all_checks(data, "config.yml", verbose = FALSE)
  })

  # Summary table
  output$summary_table <- renderTable({
    summary(report())
  })

  # Value boxes
  output$n_checks <- renderText({
    report()$n_checks_run
  })

  output$n_flags <- renderText({
    report()$n_checks_failed
  })

  # Download Excel report
  output$download_report <- downloadHandler(
    filename = function() {
      paste0("hfc_report_", Sys.Date(), ".xlsx")
    },
    content = function(file) {
      data <- read.csv(input$data_file$datapath)
      export_to_excel(report()$results, data, "hh_id", file)
    }
  )
}
```

### 11.2 Accessing individual check results

```{r shiny-individual}
# In a Shiny server, access individual results from the report
output$duplicate_details <- renderTable({
  res <- report()$results[["A01_duplicate_id"]]
  if (is.null(res)) return(data.frame(message = "Check not run"))
  data.frame(
    ID = res$flagged_ids,
    Reason = res$flag_reason,
    stringsAsFactors = FALSE
  )
})
```

## Appendix: Check Reference Table

| Code | Function | Category | Severity |
|------|----------|----------|----------|
| A.01 | `check_duplicate_ids()` | Identification | error |
| A.02 | `check_duplicate_fingerprint()` | Identification | error |
| A.03 | `check_missing_ids()` | Identification | error |
| A.04 | `check_id_format()` | Identification | error |
| A.05 | `check_id_in_sample()` | Identification | error |
| A.06 | `check_consent()` | Metadata | error |
| A.07 | `check_form_version()` | Metadata | warning |
| A.09 | `check_interview_completed()` | Metadata | warning |
| A.10 | `check_survey_tracking()` | Metadata | info |
| A.12 | `check_respondent_eligibility()` | Metadata | warning |
| B.01 | `check_enumerator_productivity()` | Enumerator | warning |
| B.02 | `check_enumerator_duration()` | Enumerator | warning |
| B.03 | `check_dk_rate()` | Enumerator | warning |
| B.08 | `check_enumerator_time_gap()` | Enumerator | warning |
| B.09 | `check_straightlining()` | Enumerator | warning |
| C.01 | `check_missing_by_variable()` | Completeness | warning/error |
| C.02 | `check_missing_by_enumerator()` | Completeness | warning |
| C.03 | `check_skip_pattern()` | Completeness | warning |
| C.04 | `check_all_missing_variables()` | Completeness | warning |
| C.07 | `check_roster_completeness()` | Completeness | warning |
| D.01 | `check_outliers_iqr()` | Outliers | warning |
| D.02 | `check_outliers_zscore()` | Outliers | warning |
| D.03 | `check_outliers_mad()` | Outliers | warning |
| D.04 | `check_hard_range()` | Outliers | error |
| D.09 | `check_negative_values()` | Outliers | warning |
| E.01 | `check_gps_boundary()` | GPS | error |
| E.02 | `check_gps_centroid_distance()` | GPS | warning |
| E.03 | `check_gps_accuracy()` | GPS | warning |
| E.04 | `check_gps_duplicates()` | GPS | warning |
| E.05 | `check_gps_swap()` | GPS | error |
| E.07 | `check_gps_null_island()` | GPS | error |
| E.08 | `check_gps_clustering()` | GPS | warning |
| E.10 | `check_gps_altitude()` | GPS | warning |
| F.01 | `check_survey_duration()` | Timing | warning |
| F.02 | `check_collection_window()` | Timing | error |
| F.03 | `check_future_dates()` | Timing | error |
| F.04 | `check_survey_end_before_start()` | Timing | error |
| F.06 | `check_duration_by_hh_size()` | Timing | warning |
| F.09 | `check_age_date_consistency()` | Logic | warning |
| G.01 | `check_hh_composition()` | Logic | warning |
| G.04 | `check_income_expenditure()` | Logic | warning |
| G.12 | `check_custom_logic()` | Logic | warning |
| M.01 | `check_benford_first_digit()` | Fabrication | warning |
| M.02 | `check_benford_second_digit()` | Fabrication | warning |
| M.03 | `check_digit_preference()` | Fabrication | warning |
| M.04 | `check_icc_enumerator()` | Fabrication | warning |
| M.06 | `check_response_entropy()` | Fabrication | warning |
| M.07 | `check_duplicate_response_patterns()` | Fabrication | error |
| M.08 | `check_variance_ratio()` | Fabrication | warning |
| O.01 | `check_other_specify()` | Text Quality | info |
| O.02 | `check_text_length()` | Text Quality | warning |
| O.03 | `check_text_duplicates()` | Text Quality | warning |
| O.04 | `check_name_format()` | Text Quality | warning |
| O.05 | `check_phone_format()` | Text Quality | warning |
